---
title: "Age and Gender Difference in Quitting and Its Rewards"
author: "Ivan Privalko"
date: "02/12/2019"
output: pdf_document
bibliography: 'C:/Users/Ivan.Privalko/Documents/ru_mobility/biblio/ru1.bib'
---


```{r, echo=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(forcats) 
library(knitr)
library(kableExtra)
library(nnet)
library(broom)


knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)

load("~/ru_mobility/genderpaper.RData")

ru2 <- ru2 %>% 
  mutate(age_groups=case_when(
    age %in% 0:35 ~  "0-35",
    age %in% 36:55 ~ "36-55",
    age %in% 56:100 ~ "56+"
  ))


```

# Abstract

The article uses data from Russia's Longitudinal Monitoring Survey to consider age and gender differences in the probability and consequences of job mobility. 
It argues that overall quit rates are similar for young men and women, but finds gender differences among older workers. 
When several personal and job characteristics are held constant, however, the quit rates of young men and women are about the same, suggesting gender differences can be explained by early career sorting. 
When controlling for individual heterogeneity (through the use of fixed effects estimation), it is found that, quitting does not improve current wages while promotions contain a significant earnings premium, at least in the short term. 
Results help to understand inequality in wages and conditions that occur due to sorting, and the importance of promotions (internal job changes with the same employer) as a "life chance" which improves earnings in the immediate sense.


# Introduction 


# Job searching and Job matching 

Two theoretical approaches make predictions about job mobility processes, these are the *job-searching* and  *job-matching* approaches. 
Both suggest that poor work environments, either from inadequate pay or inadequate conditions, push workers to search for new positions [@kalleberg2001satisfied; @gesthuizen2008mismatching]. 
The *job searching* approach considers the *Reservation Wage*, or the minimum wage a worker will consider before moving to a new position, given her skill set. 
*Reservation wage X* drives employers out of unemployment, by capturing the minimum wage that workers will accept before taking a job. 
*Reservation wage Y* where (X < Y) drives employees to new positions either within or between firms, by capturing the minimum wage that a worker will accept before changing positions. 
If a previously unemployed worker accepts *Wage X* that is below *Wage Y*, they are assumed to continue job searching until they find their reservation wage [@schmelzer2010consequences; @burdett1978theory].

The *job-matching* approach is one where workers take time to evaluate whether their skills are a good fit for their tasks at work. 
Since work is an experience good, workers must try several jobs before finding a match.
The large rates of mobility that workers see in the early stages of their career, is their navigation of the labour market, where individual employers and jobs are comapred relative to a workers skills and level of exprience [@sorensen1977structure; @thurow1975generating].  

Considering both theories together, neither explicitly notes a gender difference in job mobility. 
Despite this, authors routinely find this difference  [@blau1981race; @keith1995wage; @keith1997job; @keith1999returns]. 
Blau and Duncan [-@blau1981race] note young women are more likely to quit when compared to men. 
Keith and McWilliam [-@keith1995wage] note that economic quits happen at a simialr rate for men and women, but that family related changes are more common among young women, when comapred to young men. 
They also find that men are more likely to experience job loss, when compared to young women, suggesting that gender differences stem most from different types of mobility, rather than a gender difference in similar mobility types. 
Neither family-reated mobility nor involuntary mobility are related to processes of job matching or job searching, and so do not apply to our hypotheses directly.
However, it is important to note that women's tendency to move jobs for non-economic reasons (tied to care or other family obligations) is clearly an important piece of the gender wage-gap puzzle.
Beyond mobility itself, its effects on wages may differ by gender. For example, women may have higher reservation wages than men. Mobility may be more disruptive to their lives, especially if they have children of other family obligations which they try to balance with paid employment. As a result, they may move only for jobs with significantly higher earnings as opposed to men who are less active in the care of children and other family memebers. As a result men may see earnings as their primary responsiblility, and so incremental changes in earnings may be more lucrative for men. Cha [-@cha2014job] notes a significant difference in the effect of mobility on earnings between women with and without children, but one that runs opposite to the described pattern. Women with children saw few rewards to mobility when comapred to woment without children, whi experienced signifcant premiums in earnings. 


Although other authros find no gender differences between men and women in the wage effects of "economic quits" between young men and women in the early stages of their career [@keith1995wage].



This suggests that the wage or conditions premium described above applies more often to men than to women, even when removing gender differences in types of mobility. 
Some authors argue that the lack of a gender  


# Methodology

This article uses five rounds of Russia's Longitudinal Monitoring Survey (Rounds 20-24), a representative and longitudinal dataset of Russia's population [@kozyreva2016data]. We focus only on those who are in employment and 

# Results

Table one considers gender differences in mobility and earnings, focusing on observations rather than individuals. Overall, respondents are largely immobile although men (76 per cent) appear slightly less likely to list an observation without mobility when compared to women (81 per cent). Both men and women (4 per cent) are equally likely to list a promotion in a given year. However, promotions are generally uncommon. Respondents are much more likely to list exiting a firm, with men (18 per cent) listing more exits than women (12 per cent).

Thinking about earnings, there is an difference between men and women, with immobile observations for men (24,000+ rubles) being higher than immobile observations for women (16,000+ rubles). However, between each group, there appears to be a premium tied to promotion, and a penalty tied to Exit. 

```{r}

ru2 %>% 
  filter(!is.na(mob_final)) %>% 
  group_by(gender, mob_final) %>% 
  summarise(n = n(),
            wage = mean(wage, na.rm=TRUE)) %>% 
  ungroup() %>% 
  group_by(gender) %>% 
  mutate(per = round(n/sum(n)*(100), 2)) %>% 
  ungroup() %>% 
  select(-gender,
         Mobility= mob_final,
         Observations = n,
         Percent = per,
         Wage= wage) %>% 
  kable(format.args = list(decimal.mark = '.', big.mark = ","),
        caption = "Mobility and earnings, by gender") %>% 
  kable_styling(latex_options = 'hold_position') %>% 
  group_rows("Male", 1, 4) %>% 
  group_rows("Female", 5, 8) 
```

Table 2 considers the age differences in these rates. We see that younger groups are more mobile than older groups for both genders, with a seemingly steady increase between respondents aged 0-35, 36-55, and those aged over 55. Young men are still more mobile than young women, but the likelihood of mobility decreases among older respondents.  

```{r}

ru2 %>% 
  filter(!is.na(mob_final)) %>% 
  group_by(gender, age_groups) %>% 
  count(mob_final) %>% 
  mutate(per = round(n/sum(n)*(100), 2)) %>% 
  ungroup() %>% 
  select(Mobility= mob_final,
         Observations = n,
         Percent = per) %>% 
  kable(format.args = list(decimal.mark = '.', big.mark = ","),
        caption = "Mobility, by age and gender") %>% 
  kable_styling(latex_options = 'hold_position') %>% 
  group_rows("Male Aged 0-35", 1, 4) %>% 
  group_rows("Male 36-55", 5, 8) %>% 
  group_rows("Male 56+", 9, 12) %>% 
  group_rows("Female Aged 0-35", 13, 16) %>% 
  group_rows("Female Aged 36-55", 17, 20) %>% 
  group_rows("Female Aged 56+", 21, 24)
```

## Multinomial regression

this section cosiders the likelihood of experiencing a mobility event.

```{r}
d1  <-  ru2 %>% 
  filter(round == 24,
         age_groups=='0-35',
         gender=='Male') 


  
m1  <- multinom(mob_final ~ age + gender + hours, data = d1)
```


```{r}

summary(m1)
coef(m1)
tidy(m1, conf.int = T,
     conf.level = 0.95,
     exponentiate = T) %>% 
  ggplot(aes(x=term, y=estimate))+
  geom_point(aes(colour=term))+
  facet_wrap(~y.level)+
  geom_errorbarh(aes(xmin=estimate-std.error,
                    xmax=estimate+std.error,
                    colour=term))+
  coord_flip()

```


# References


