---
title: "Age and Gender Differences in Job Mobility and Its Rewards"
author: "Ivan Privalko"
date: "02 December 2019"
output: pdf_document
bibliography: 'C:/Users/Ivan/Desktop/R projects/ru_mobility/biblio/ru1.bib'
toc: TRUE
---


```{r, echo=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(forcats) 
library(knitr)
library(kableExtra)
library(nnet)
library(broom)


knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)

load("C:/Users/Ivan/Desktop/R projects/ru_mobility/genderpaper.RData")

ru2 <- ru2 %>% 
  mutate(age_groups=case_when(
    age %in% 0:35 ~  "0-35",
    age %in% 36:55 ~ "36-55",
    age %in% 56:100 ~ "56+"
  ))


```

# Abstract

The article uses data from Russia's Longitudinal Monitoring Survey to consider age and gender differences in the probability and consequences of quits and promotions. 
It argues that overall quit rates are similar for young men and women, but finds gender differences among older workers. 
When several personal and job characteristics are held constant, however, the quit rates of young men and women are about the same, suggesting gender differences are explained by early career sorting. 
When controlling for individual heterogeneity (through the use of fixed effects estimation), it is found that, quitting does not improve current wages while promotions contain a significant earnings premium, at least in the short term. 
Results help to understand inequality in wages and conditions that occur due to sorting, and the importance of promotions (internal job changes with the same employer) as a "life chance" which improves earnings in the immediate sense.


# Introduction 


# Job searching and Job matching 

Two theoretical approaches make predictions about job mobility, these are the *job-searching* and  *job-matching* approaches. 
Both suggest that poor work environments, either from inadequate pay or inadequate conditions, push workers to search for new positions [@kalleberg2001satisfied; @gesthuizen2008mismatching]. 
The *job searching* approach considers the *Reservation Wage*, or the minimum wage a worker will consider before moving to a new position, given her skill set. 
*Reservation wage X* drives employers out of unemployment, by capturing the minimum wage that workers will accept before taking a job. 
*Reservation wage Y* where (X < Y) drives employees to new positions either within or between firms, by capturing the minimum wage that a worker will accept before changing positions. 
If a previously unemployed worker accepts *Wage X* that is below *Wage Y*, they are assumed to continue job searching until they find their reservation wage [@schmelzer2010consequences; @burdett1978theory].

The *job-matching* approach is one where workers take time to evaluate whether their skills are a good fit for their tasks at work. 
Since work is an experience good, workers must try several jobs before finding a match.
The large rates of mobility that workers see in the early stages of their career, is their navigation of the labour market, where individual employers and jobs are comapred relative to a workers skills and level of exprience [@sorensen1977structure; @thurow1975generating]. 

How do the theories above explains the large differencs in mobility between workers? 
Both understand that workers are most mobile in the early stages of their career. 
However, in the job searching approach, individuals try to secure their reservation wages as they establish themselves in the labour market and gain the experience and networks needed to evaluate new positions. 
As they gradually close the gap between their given wage and their reservation wage, they become less mobile [@sorensen1977structure; @kalleberg1979sociology; @sorensen1975structure].
The job-matching appraoch is slightly different in that it does not expect a wage premium tied to mobility, only that the subjectively evaluated match between a worker and her position should improve with mobility. 
This approach also anticipates the large rate of mobility in the early stages of a person's career, but in this mechanism workers are "experiencing" the type of work they would like to do. 
It is possible that this mechanism affects men and women differently, in that men would be more driven by wages, and women be more driven by work life balance, especially if women hold the majority of care responsibilities in the home [@keith1997job].

# Gender differences in mobility and its returns

Considering both theories together, neither explicitly notes a gender difference in job mobility. 
Despite this, authors routinely find this difference  [@blau1981race; @keith1995wage; @keith1997job; @keith1999returns]. 
Blau and Duncan [-@blau1981race] note young women are more likely to quit when compared to men. 
Keith and McWilliam [-@keith1995wage] note that economic quits happen at a simialr rate for men and women, but that family related changes are more common among young women, when comapred to young men. 
They also find that men are more likely to experience job loss, when compared to young women, suggesting that gender differences stem mostly from different types of mobility, rather than a gender difference in wage returns from similar mobility types [@keith1995wage; @keith1999returns].
In short, part of the voluntary job mobility carried out by women is closer tied to job matching than job searching, in this way Keith and McWilliams [-@keith1995wage; -@keith1999returns] sugegst women find positions where they are better able to combine family roles and work. 
Gesthuizen finds some evidence to support this, reporting that dissatisfaction with the job match and dissatisfaction with hours increases women's chances of voluntary mobility to a new employer.
This effect does not emerge for men. 
Further, subjective dissatisfaction with wages also pushes women to inter-firm mobility, but not men, which would support the job searching hypothesis. 
In discussing the result Gesthuizen [] notes that mobility may be “less costly” for women, in that men are more likely typically “breadwinner” roles, with less room
Keith and McWilliams [-@keith1999returns] also note a gender difference in job searching behaviour, reporting men are more likely to explore new positions than women. Job searching behaviour was particualrly important as it brought a significant wage premiums for both men and women who engaged in job mobility. 

Beyond gender differences in mobility types, mobility may have different effects on men and women. 
Keith and McWilliams [-@keith1999returns, -@keith1995wage; -@keith1997job] find no gender difference in the effects of job mobility on wages, assuming the mobility occurs for economic reasons. 
Further, they find that men and women benefit roughly equally from the interaction between job search and job mobility.
Gesthuizen [] also finds no significant difference between men and women in the premium tied to mobility.
Cha [-@cha2014job] notes a significant difference in the effect of mobility on earnings between women with and without children, but one that runs opposite to the described pattern. 
Women with children see few rewards to mobility when compared to woment without children, who experienced signifcant premiums in earnings.
The study doesn't consider men.
Kronberg [-@kronberg2013stay] finds this specific difference, reporting that men gain more from voluntary mobility to a new employer when comapred to women, although see a premium also. 


Although other authros find no gender differences between men and women in the wage effects of "economic quits" between young men and women in the early stages of their career [@keith1995wage].



This suggests that the wage or conditions premium described above applies more often to men than to women, even when removing gender differences in types of mobility. 
Some authors argue that the lack of a gender  


# Methodology

This article uses five rounds of Russia's Longitudinal Monitoring Survey (Rounds 20-24), a representative and longitudinal dataset of Russia's population [@kozyreva2016data]. We focus only on those who are in employment and 

# Results

This section is split into three pars. 
First, I present gender differences in mobility, and the wages tied to certain mobility responses.
Second, I explore gender differences in mobility using multinomial logistic regression.
Finally, we explore the effects of mobility on earnings using fixed-effects linear regression.

## Descriptive statistics

Figure 1  considers gender differences in mobility types focusing on observations rather than individuals. 
Overall, respondents are largely immobile although women (0.814) are less mobile compared to men (0.762). 
Both men and women (0.044) are equally likely to list a promotion in a given year, although promotions are uncommon in general. 
Respondents are much more likely to list exiting a firm, with men (0.182) listing more exits than women (0.130).
Lateral changes within the firm are the least common and do not appear to have a gender difference. 
Considering the confidence intervals in Figure 1, it seems gender differences are most prominent in terms of firm exits.
Unfortunately we are not able to discern voluntary from involutnary exits, which are particualrly important for discussions of reservations wages.
However, this category, along with promotions, likely contains the job mobility which is driven by reservation wages.
 

```{r, fig.cap="Proportion of respondents citing mobility type by gender"}

sum1 <- ru2 %>%
filter(!is.na(mob_final)) %>% 
  group_by(gender, mob_final) %>%
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n),
         prop_dev =prop*(1-prop),
         se=sqrt(prop_dev/n),
         conf_up = prop + (se*1.96),
         conf_low = prop -(se*1.96))  
  
right_label <- sum1 %>%
  filter(gender=='Male')

left_label <- sum1 %>% 
  filter(gender=="Female")

sum1 %>% 
ggplot(aes(x=mob_final,
             y=prop,
             colour=gender))+
  geom_point(aes(size=0.3))+
  geom_errorbar(aes(ymin=conf_up,
                    ymax=conf_low),
                width=0.3,
                size=0.3)+
  labs(subtitle= "Estimates contain 95% confidence intervals (Standard error*1.96)",
       x="Mobility type",
       y="Proportion of respondents",
       caption="RLMS rounds 2015 to 2011 \nAuthors own calculation")+
  guides(size=FALSE)+
  geom_text(data = right_label, aes(color = gender, label = round(prop, 3)),
                               size = 3, hjust = -.6) +
  geom_text(data = left_label, aes(color = gender, label = round(prop, 3)),
            size = 3, hjust = 1.8)

```


Figure 2 considers the age differences tied to the changes in Figure 1. We note that younger respondents are more likely to experience promotions and exits. Both of these measures gradually decline as we consider older groups. Noteably, gender differences in quits dissapear for older respondents, but remain for younger respondents. This effect stems from the sharp decline in men's likelihood of quitting. 

```{r, fig.cap="Age and gender differences in mobility"}

sum2 <- ru2 %>%
filter(!is.na(mob_final)) %>% 
  group_by(gender, age_groups , mob_final) %>%
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n),
         prop_dev =prop*(1-prop),
         se=sqrt(prop_dev/n),
         conf_up = prop + (se*1.96),
         conf_low = prop -(se*1.96))  
  
right_label <- sum1 %>%
  filter(gender=='Male')

left_label <- sum1 %>% 
  filter(gender=="Female")

sum2 %>% 
ggplot(aes(x=mob_final,
             y=prop,
             colour=gender))+
  geom_point(aes(size=0.3))+
  geom_errorbar(aes(ymin=conf_up,
                    ymax=conf_low),
                size=0.3)+
  labs(subtitle= "Estimates contain 95% confidence intervals (Standard error*1.96)",
       x="Mobility type",
       y="Proportion of respondents",
       caption="RLMS rounds 2015 to 2011 \nAuthors own calculation")+
  guides(size=FALSE)+
  facet_wrap(~age_groups)+
  scale_x_discrete(labels = abbreviate)

```

Generally, there is a gender and age effect tied job mobility, with younger workers being more mobile than older workers, and younger men being more mobile than younger women. We now turn to the wage differences between these groups. Importantly, we will not consider the effect of promotion or exit on wages, focusing instead on the average wage associated with a given gender and a given mobility type. 



```{r, fig.cap="Wage differences between gender and mobility groups"}
sum3 <- ru2 %>%
  filter(!is.na(mob_final),
         !is.na(wage)) %>% 
  group_by(gender, mob_final) %>% 
  summarise(wage_mean = mean(wage, na.rm=TRUE),
            wage_sd = sd(wage, na.rm = TRUE),
            n= n(),
            wage_se = wage_sd/sqrt(n)) 
 

sum3 %>% 
ggplot(aes(x=mob_final, 
             y= wage_mean,
             col=gender))+
  geom_line(aes(group=gender))+
  geom_point(aes(group=mob_final,
                 size=3))+
  geom_errorbar(aes(ymin=wage_mean-(wage_se*1.96),
                    ymax=wage_mean+(wage_se*1.96)),
                width=0.25)+
  scale_y_continuous(labels = scales::comma)+
  labs(subtitle= "Estimates contain 95% confidence intervals (Standard error*1.96)",
       y= "Average monthly wages in rubles",
       x="Mobility type",
       caption="RLMS rounds 2015 to 2011 \nAuthors own calculation")+
  guides(size=FALSE)+
  geom_text(data = sum3, aes(color = gender, label = round(wage_mean, 0)),
            size = 3, hjust = -.6)


```

Figure 4 captures thre eimportant differences. First, there is a gender difference tied to pay, with men earning more than women in each mobility cateogry. Second, there is a bility difference in pay, with respondents who experience a promotion in the survey year citing significantly more in terms of average earnings, compared to respondents citing a different mobility type in a given survey year. Third, there is a pay penatly among respondents who cite employment exit, but this difference is only significant for men. This effect likely stems from women's pay reaching a "floor". We consider the age differences in this effect below.


```{r, fig.cap="Wage differences between age, gender, and mobility groups"}
sum4 <- ru2 %>%
  filter(!is.na(mob_final),
         !is.na(wage)) %>% 
  group_by(gender, age_groups, mob_final) %>% 
  summarise(wage_mean = mean(wage, na.rm=TRUE),
            wage_sd = sd(wage, na.rm = TRUE),
            n= n(),
            wage_se = wage_sd/sqrt(n)) 
 

sum4 %>% 
ggplot(aes(x=mob_final, 
             y= wage_mean,
             col=gender))+
  geom_line(aes(group=gender))+
  geom_point(aes(group=mob_final,
                 size=3))+
  geom_errorbar(aes(ymin=wage_mean-(wage_se*1.96),
                    ymax=wage_mean+(wage_se*1.96)),
                width=0.25)+
  scale_y_continuous(labels = scales::comma)+
  labs(subtitle= "Estimates contain 95% confidence intervals (Standard error*1.96)",
       y= "Average monthly wages in rubles",
       x="Mobility type",
       caption="RLMS rounds 2015 to 2011 \nAuthors own calculation")+
  guides(size=FALSE)+
  facet_wrap(~age_groups)+
  scale_x_discrete(labels = abbreviate)


```

Thinking of the patterns in Figure 4, respondents appear to gain the most from mobility aged 36-55, although the premium only appears to apply to men. Surprisingly, older groups report the lowest income and also the lowest return on mobilty. Older women appear to make significant gainst when citing promotion, however the large confidence intervals suggest this premium is not significantly different from other age groups.

## Multinomial logistic regression

We now consider some explanatory measures tied tojob mobility differences. 


```{r}
d1<- ru2 %>% 
  filter(round==24)

m1 <- multinom(mob_final ~ gender + age_groups + tenure , data=ru2)

kable(tidy(m1), digits = 3)

```


# References


