---
title: "Gender Differences in Russia's Job Mobility and Its Rewards"
author: "Author"
date: "02 December 2019"
output:
  pdf_document: default
  word_document: default
bibliography: C:/Users/Ivan.Privalko/Desktop/R projects/ru_mobility/biblio/ru1.bib
toc: yes
always_allow_html: true
---


```{r, echo=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(forcats) 
library(gtools)
library(knitr)
library(kableExtra)
library(nnet)
library(broom)
library(plm)


knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)

load("C:/Users/Ivan.Privalko/Desktop/R projects/ru_mobility/genderpaper.RData")

ru3 <- ru3 %>% 
  mutate(age_groups=case_when(
    age %in% 0:35 ~  "0-35",
    age %in% 36:55 ~ "36-55",
    age %in% 56:100 ~ "56+")) %>% 
  filter(hours > 0)


ru3 %>% 
  filter(econ=='Employed') 


```
\newpage
# Abstract

This article uses data from Russia's Longitudinal Monitoring Survey to consider age and gender differences in the probability and consequences of firm exits and promotions.
Russia's labour market should have high rates of job mobility.
However,  IMF figures suggest that mobility in the early 2010's is on par with mobility recorded in the 1980's.
Beyond this, little has been said on who is mobile and whether mobility has any impact on wages, once personal characteristics of movers are controlled for.
In other words, little is known about whether job mobility is a sorting mechanism, or whether it has premiums in terms of wages in itself.
Results show significant differences between age and gender groups regarding exits and promotions.
On promotions, when several personal and job characteristics are held constant, instances of promotion for young men and women are about the same.
However, the wage benefits tied to promotion are particualrly important for young women when comapred to yougn men.
On the topic of firm exits, when several personal and job characteristics are held constatnt, job exit was more common among younge men than young women.
This was also true of middle aged men and women.
However, young women saw no effect in wages after a firm exit, while young men saw a significant decrease in subsequent wages following an exit. 
These results are flipped for middle aged workers, as men saw no change in wages following an exit, but women saw a significant decrease in subsequent wages following an exit.
In general, the early stages of a respondent's career are marked by periods of high mobility in Russia, which is similar to the experience of young workers in other countries.
After this period, mobility becomes increasingly unlikely.
Part of this result could stem from the premiums tied to promotion.
Results help to understand processes of inequality in wages and conditions that occur due to sorting, and the importance of promotions (internal job changes with the same employer) as "life chances" which improve earnings in the immediate sense.
Gender differences in securing these life chances may help to understand wider gender gaps in earnings, which emerge later. 

# Introduction 

Russia's liberalised economy gave workers more employment options and a wider spectrum of potential wages and conditions [@imf1991study; @clarke2002market; @clarke2000closure; @kozyreva2015economic], at least when comapred to the Soviet Union. 
Theorists assumed that workers would increase their job mobility in order to correct earnings and conditions [@keith1995wage; @keith1997job; @cha2014job; @clarke2002market].
Such corrections are an important part of *job searching* and *job matching* theories of inequality [@schmelzer2012consequences; @schmelzer2011consequences].
Despite these predictions, Russian job mobility in 2010 was lower than the rate in 1980's Soviet Union [@imf1991study]. 
Despite seeing significant churn immediately after the reforms of the 1990's, the job mobility rate gradually declined and returned to previous levels, with workers preferring to remain in the same job with the same employer [@clarke2000closure]. For this reason Russia presents an interesting puzzle for researchers, one where wage inequality exists parallel to low job mobility.

This article has two aims. 
First, we will focus on predicting mobility and estimating group differences in who moves and who remains in the same job with the same employer. 
Do gender differences in Russia's job mobility exist, and if they do are these the product of differences in occupational sorting or tenure, or marital status? 
Further, is mobility confined only to younger workers, or do certain occupations contain mobility patterns that cannot be explained by age differences between workers alone?
Second, having established these differences, we  consider the effects of job mobility on wages. 
Does mobility reward workers in terms of wages, or is job mobility simply a sorting mechanisms used by workers with specific characteristics?

We will use the Russia Longitudinal Monitoring Survey (RLMS) throughout  as a representative sample of the country's workforce. 
This approach has two strengths which are relevant to the wider discussion of job mobility. 
First, the survey explicitly measures firm exits and promotions which occurred in the last twelve months.
Previous authors discussing mobility often operationalise promotions and exits in complex ways [@kalleberg2001satisfied; @keith1995wage; @keith1997job].
Second, using longitudinal data we can focus on the effects of mobility on wages, instead of comparing the wages of movers to non-movers [@cha2014job; @reichelt2017occupational]. Since mobility is not a random event, we will remove the influence of unmeasured person-specific characteristics and consider the net gains produced by job mobility. 

The article has three findings. First, mobility is most likely in the earlies stages of one's career. 
For young and middle aged workers, age and firm tenure are negatively correlated with mobility. 
This is true of both promotions and firm exits and suggests that young workers must be mobile in the early stages of their careers', before slowing down their pattern of mobility, as in many other countries.
Second, there is no difference between young women and young men in terms of promotion after controlling for a range of factors like working time and marital status.
This is also true of middle-aged workers, who see no gender difference in promotion after we control for a range of factors.
A gender difference exists in terms of firm exits, with young men being more likely to exit a firm when compared to young women, but no other gender difference exists.
Finally, there may be no gender difference in the likelihood of promotion between young men and young women, but we will find a singificant difference in the returns to promotion when comparing young men to young women, with young women benefitting more from the change compared to young men.
Further we find a gender difference in the effects of firm exit among young men and women, with young women being unaffected by firm exit and young men seeing a signifciant fall in wages after the change.
Gender differences in returns to mobility do not exist forr middle aged workers.

The article is structured as follows. 
The first section summarises the theoretical literature on job mobility. 
The second section summarises the empirical findings between age, gender, and job mobility. 
This section also presents Russia as an important case for studies of job mobility. 
The third section summarises our data and approach of estimating mobility, and estimating the effects of mobility. 
Section four presents the results, discusses the findgings and concludes the article. 

# Review

## Job searching and Job matching 

Studies of job mobility rely on two theoretical approaches, *job-searching* and  *job-matching*. 
Both suggest that poor work environments either from inadequate pay or inadequate conditions, push workers to search for new positions [@kalleberg2001satisfied; @gesthuizen2008mismatching; @schmelzer2011consequences]. 
The *job searching* approach considers the *Reservation Wage*. 
This is the minimum wage a worker accepts before moving to a new position, given their skill set. 
*Reservation wage X* draws employees out of unemployment and is the minimum wage that he or she will accept before taking a job.
*Reservation wage Y* where (X < Y) draws employees to new positions either within or between firms; it is the minimum wage that a worker will accept before changing positions.
If a previously unemployed worker accepts *Wage X* that is below *Wage Y*, they are assumed to continue searching until they secure *Wage Y* [@schmelzer2010consequences; @burdett1978theory].
*Wage Y* is typically greater than *Wage X* for several reasons including the fact that job-searching is costly to workers. 
Most importantly however, workers searching for new positions have access to wider networks of information and wider knowledge about possible reservation wages. 
There may be gender differences in these networks, which can lead to differences in reservation wages after the change.

On the other hand, the *job-matching* approach is one where workers take time to evaluate whether they have a good amtch between their skills and their tasks at work. 
Since work is an experience good, workers must try several jobs before finding a match.
The large rates of mobility that workers see in the early stages of their career, is them navigating the labour market, trying individual employers and jobs and comparing their experience with rpevious experiences [@sorensen1977structure; @thurow1975generating]. 

How do the theories above explains the large differences in mobility between workers? 
Both understand that workers are most mobile in the early stages of their career. 
However, in the job searching approach, individuals try to secure their reservation wages as they establish themselves in the labour market and gain the experience and networks needed to evaluate new positions. 
As they gradually close the gap between their given wage and their reservation wage, they become less mobile [@sorensen1977structure; @kalleberg1979sociology; @sorensen1975structure].
The job-matching approach is slightly different in that it does not expect a wage premium tied to mobility, only that the subjectively evaluated match between a worker and her position should improve following mobility. 
This approach also anticipates the large rate of mobility in the early stages of a person's career, but in this mechanism workers are "experiencing" the type of work they would like to do. 
It is possible that this mechanism affects men and women differently, in that men would be more driven by wages, and women be more driven by work life balance or their experience at work. This is likely especially true if women hold the majority of care responsibilities in the home [@keith1997job].

## Gender differences in mobility and its returns

### Gender and mobility

Considering both theories together, neither explicitly notes a gender difference in job mobility. 
In the empirical literature too, the findings are generally mixed, even when considering young workers. 
Blau and Duncan [-@blau1981race] find that young women are more likely to quit compared to young men in a sample of US workers. 
Using more recent data, Hachen [-@hachen1988gender] finds women are less likely to exit, but are even less likely to receive promotion relative to men, even when controlling for a range of measures including share of women in an occupation.

Gesthuizen [-@gesthuizen2009job] finds that women are more likely to exit the firm and more likely to secure promotion. compared to men, but that measures of workplace dissatisfaction explain both of these differences.
In these results it seems that dissatisfaction is a stronger predictor of exit for women, when compared to men and that dissatisfaction with working hours and dissatisfaction with the job are stronger predictors of promotion for women when compared to men.
Keith and McWilliam [-@keith1995wage] do not find a gender difference when predicting "economic quits". 
However, they find that "family-related quits" are more common among young women, compared to young men, and that involuntary exit was more common among men that it was among women.
Keith and McWilliams [-@keith1999returns] also note a gender difference in job searching behaviour, reporting men are more likely to explore new positions than women. 
Job searching behaviour was particularly important as it brought a significant wage premiums for both men and women who changed jobs. 
In general, results tend to suggest that differences in mobility between men and women can be explained by other factors like workplace dissatisfaction.

### Gender and returns to mobility

The theories above also do not consider gender differences in the returns to mobility.
Here too, authors have produced mixed findings, with some supporting and some refuting gender differences in returns to mobility.
Keith and McWilliams [-@keith1999returns; -@keith1997job] reports no gender difference in the effects of job mobility on wages, assuming the mobility occurs for "economic reasons". 
Gesthuizen [@gesthuizen2008mismatching] also finds no significant difference between men and women in subjective premiums tied to mobility within and between firms.
Kronberg [-@kronberg2013stay] finds a gender difference in mobility's effect on earnings using US data. 
She notes that men gain more from a voluntary exit to a new employer when compared to women, although women see a premium also. 
Crucially Kronberg [-@kronberg2013stay] notes that the effect applies only to men and women in "good jobs", characterised as those in high paying occupations with benefits and insurance.

Part of this effect may also stem from family types and other household characteristics. 
Cha [-@cha2014job] reports a difference between single women and women with children regarding the effect of a quit on earnings. 
According to Cha [-@cha2014job] women with children see almost no rewards after the change, compared to women without children who see significant premiums.
However, since the study does not consider men, it is not possible to comment on gender differences in quit and promotion premiums.
Fuller [-@fuller2008job] considers the effect of mobility and wages for both men and women, finding that early quits  during the first five years of one's career yield an equally positive result for both men and women.
This result also emerges for early career movers with children, both men and women saw a positive increase in earnings after the change and no gender difference was detected.
When Fuller [@fuller2008job] substitutes marital status for children a similar result emerges, men and women see similar returns to job mobility regardless of their marital status. 
In short, results suggest some premium tied to volutnary mobility exists, however, this premium appears equally likely for both men and women, according to the literature.

### Gender and Russia's labour market, as a case

Most of the papers above rely on US data, and so may not apply to Russia easily.
Gerber and Mayorova [-@gerber2006dynamic] for example, contradict many of the findings above, reporting that throughout the 1990's women relative to men, had higher rates of layoff and labour market exit, and higher odds of starting in low-quality jobs. 
Further, they find that women reported lower odds of voluntary job mobility overall, and lower odds of holding good quality jobs.
In short, while Russia's transition to a market economy may have improved women's access to jobs, women remain at a disadavatnage relative to men [@gerber2006dynamic].

Separately, Clarke [-@clarke2002market] notes the sharp rise in wage differences most of which which is located within occupations, regions, and similar positions.
Clarke [-@clarke2000closure] also notes that despite the high rates of wage inequality and job mobility in the 1990's, job mobility has declined without a corresponding fall in wage differences.
He hypothesises that Russia is a country where *"there are significant barrieirs to labour mobility that are preventing the erosion of these differentials"* [in wages].
Again, this large inequality in wages is particularly strange because it occurs within occupations and within regions, suggesting that workers have much to gain from local job mobility, if only in terms of wages and not other factors like housing or access to employer sponsored childcare.
Clarke concludes [-@clarke2000closure] that personal acquintances and individual networks are particualrly important in securing high wages in the wider labour market, this mechanism has led to a "closing" of the Russian labour market since the 1990's, making Russia a particularly interesting case for study.

# Methodology

This article uses five rounds of Russia's Longitudinal Monitoring Survey (Rounds 20-24) as a representative and longitudinal sample of Russia's population [@kozyreva2015economic]. 
This approach covers years 2011 to 2015 and avoids the European Debt Crisis, but partially includes the Russian financial crisis between 2014 and 2015. 
Using this sample we draw on objectrive measures of earnings and job mobility for men and women, we also use respondents age to consider braod age categories for young workers, middle aged workers, and older workers.
We will consider two estimation techniques for both of the aims stated above. 
When predicting group differences in job mobility, we consider multinomial logistic regression with clustered standard errors.
When estimating the effects of different types of mobility on earnings, we will use fixed-effects linear models. 
We will discuss each of these details in the subsections below.

## Sample

The RLMS in a longitudinal study of Russian respondents [@kozyreva2015economic; @kozyreva2016data]. 
Data is collected at the household level annualy.
The survey contains a rich range of questions on employment and earnings, inluding simple measures of job mobility over the last twelve months.
The survey is particualrly useful to this article because both internal and external job changes are captured by the survey.
I define the sample as; observations where respondents are in employment throughout the five waves; observations where respondents are not in self-employment, education, or inactivity; and observations where respondents have no missing values for job history and mobility.
The final data frame takes the form of a person-year file, which ignores households and focuses on individual responses.
It contains 15,700 respondents and 42,400+ observations.

## Variables

This section considers three different sets of variables; job mobility, wages, and a standard set of controls mentioned in the literature review above. 
The article's definition of job mobility relies on three variables within the RLMS dataset. The first asks respondents whether they changed jobs since November of the previous year (IXNEWJOB).
Using this measure we can capture the most basic form of mobility, whether respondents remain in the same job with the same employer, whether they changed jobs with their employer, and whehter they changed employers. 
The next measure asks respondents whether they received a promotion since November of the previous year (IXPROMOT), allowing us to separate promotions from other internal moves, such as lateral changes.
The final measure asks respondents whether they experienced a lateral move since November of the previous year (IXMOVOA), which allows us to capture laterla changes which occur within the same employer, but which are not promotions.
As a result we are left with a four category measure of mobility, capturing respondents in the same job with the same employer, respondents who exit the firm, respondents who receive a promotion, and respondents who experience a lateral move within the same firm. 
One limitation of this measure is that we cannot distinguish between volutnary and involuntary firm exits. 
Thus we must assume that the majority of exits from a given employer are voluntary changes to a new employer.

Regardng wages, we consider monthly net wages from the respondent's primary job (IXWAGELM). This considers monthly earnings in rubles; if respondents are paid in a different currency they are asked to convert the monthly amount to rubles. 
We specifically focus on wages from the main job; and ignore other measures like total monthly income (IXINCLMO). 
This is to capture the reservation wage mentioned above. 
Income from other sources like returns on savings, or returns on rental properties, do not encompass a reservation wage, and should be ignored.

Finally, we consider a set of controls for models predicting mobility and models predicting wages.
The literature above notes that russian wages tend to change quickly over time. 
Although the majority of authors above mention the Russian labour amrket of the 1990's, it is clear that wages, in an aggregate sense, tend to change over time. 
As a result we control for annual changes using survey round dummies (ROUND).
Beyond this, authors mention the importance of human capital and experience; while we cannot measure a respondents full labour market experience directly, we can capture some of this effect using a respondent's age and age squared (IXAGE).
Further, we note the importance of tenure as a measure of firm specific human capital, or the level of internal labour market experience reported by respondents.
We capture this measure using the year the respondent started with their current employer (IXJBSYR), and the current survey year.
We also want to capture "pure" mobility effects [@le2002job], and so we controls for occupational differences (IXILPJB8) in mobility as well as the effects of occupational mobility on wages.
By "pure mobility effect" we mean the change of jobs between and within firms while remaining in the same occupation [@le2002job]. 
Finally we consider whether respondents hold supervisory positions within the same occupation (IXPRISUB), in an effort to avoid changes in responsibility within the same role.

## Estimation

We will consider two sets of models throughout the article. 
The first predicts the likelihood of promotion or firm exit, relative to staying in the same job with the same employer. 
The second estimates the effects of job mobility on the log wages of respondents. 

We first consider predictions which focus on mobility events, using logistic regression. We can start by thinking about a simple binary regression which considers the likelihood of a mobility event, like a quit $(y_i = 1 | x_i)$, compared to it not occuring, like staying in the same job with the same employer $(y_i = 0 | x_i)$. This model is shown below.

$$Pr(y_i = 1 | x_i) = \frac{exp(\beta_0 + \beta_1 x_i)}{1 + exp(\beta_0 + \beta_1x_i)}$$

This model depends on an intercept $(\beta_0)$, a set of estimated effects $(\beta_n)$, and a set of predictors $(x_n)$. Rabe and Hesketh [-@rabe2008multilevel] show how this model can be extended to include more than one outcome. For example, if the number of outcomes is three $(S=3)$, they propose the following model.

$$Pr(y_i = 1 | x_i) = \frac{1}{1 + exp(\beta_0^{[2]} + \beta_1^{[2]}x_i) + exp(\beta_0^{[3]} + \beta_1^{[3]}x_i)} $$
$$Pr(y_i = 2 | x_i) = \frac{exp(\beta_0^{[2]} + \beta_1^{[2]}x_i)}{1 + exp(\beta_0^{[2]} + \beta_1^{[2]}x_i) + exp(\beta_0^{[3]} + \beta_1^{[3]}x_i)} $$
$$Pr(y_i = 3 | x_i) = \frac{exp(\beta_0^{[3]} + \beta_1^{[3]}x_i)}{1 + exp(\beta_0^{[2]} + \beta_1^{[2]}x_i) + exp(\beta_0^{[3]} + \beta_1^{[3]}x_i)} $$

In the binary case we rely only on one intercept $(\beta_0)$ and one estimated coefficient $(\beta_1)$. 
With multiple outcomes we rely on an intercept for each outcome $(\beta_0^{[1]}, \beta_0^{[2]})$, and an estimated coefficient for each outcome $(\beta_1^{[1]}, \beta_1^{[2]})$.

In order to estimate the predicted effect of moving on  wages we will turn to fixed effects linear regression.
We can think of a model that predicts wages as having two error-terms, one specific to individual $i$ which does not vary over time $t$, and one which varies over $i$ and $t$ [@longhi2014practical; @rabe2008multilevel].
The resulting model is listed below.

$$y_{it}^* = \beta_1 X_{1it} +\beta_2 X_{2it} + ... + \beta_n X_{nit} + \alpha_i + u_{it}$$
The resulting model predicts wages $(y_{it})$ for individual $i$ at time $t$, given the estimated effect $(\beta_n)$ of each control $(x_{nit})$, and two error terms; one varying between individuals $(\alpha_i)$, and one varying between and within individuals $(u_{it})$.
However, we are interested mostly in the effects of mobility within individuals. 
Since mobility is not a random process personal, time-fixed, error terms of respondents $(\alpha_i)$, things like motivation, upbringing, or self-sufficiency for example, will likely correlate with a person's chance of exiting a firm or taking a promotion.
We can remove their inluence on mobility using a "within-estimator", by subtracting each terms from its cluster mean [@longhi2014practical; @rabe2008multilevel; @allison2009fixed].
This formula is listed below.

$$y_{it} - \overline{y_i} = \beta(X_{it} - \overline{X_i}) + (\alpha_i - \overline{\alpha_i}) + (u_{it} - \overline{u_i})$$

This approach models average deviation in wages $(y_{it} - \overline{y_i})$ for individual $i$ at a given time $t$.
The model also considers person specific deviations from a set of controls, which include job mobility $\beta(X_{it} - \overline{X_i})$. 
Importantly, this model omits person-specific errors and avoids individual heterogeneity tied to job mobility $(\alpha_i - \overline{\alpha_i})$.
The remaining level-2 errors are tied to differences over time that are not person-specific $(u_{it} - \overline{u_i})$.

# Results

This section is split into three parts. 
First, we will discuss summary statistics which look at gender differences in mobility, and gender differences in wages by movers and non-movers.
Second, we estimate gender differences in mobility using multinomial logistic regression.
Finally, we will explore the effects of mobility on earnings using fixed-effects linear regression, as mentioned above.

## Descriptive statistics

Figure 1  considers gender differences in mobility types focusing on observations rather than individuals.
We are primarily interested in whether men and women move at a similar rate using observations alone.
Overall, respondents are largely immobile although women (0.814) are less mobile compared to men (0.762). 
This difference mainly stems from men's higher likelihood of listing a firm exit, when comapred to women.
Both men and women (0.044) are equally likely to list a promotion in a given year, although promotions are uncommon in general, if we consider cross-sectional data alone. Respondents are much more likely to list exiting a firm, with men (0.182) listing more exits than women (0.130).
Lateral changes within the firm are the least common and do not appear to have a gender difference. 
Considering the confidence intervals in Figure 1, it seems gender differences are most prominent in terms of firm exits.
Unfortunately we are not able to discern between voluntary and involuntary exits, which are particualrly important for discussions of reservations wages.
However, this category, along with promotions, likely contains a majority of voluntary exits, driven by reservation wages.
 

```{r, fig.cap="Proportion of respondents citing mobility type by gender"}

sum1 <- ru3 %>%
filter(!is.na(mob_final)) %>% 
  group_by(gender, mob_final) %>%
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n),
         prop_dev =prop*(1-prop),
         se=sqrt(prop_dev/n),
         conf_up = prop + (se*1.96),
         conf_low = prop -(se*1.96))  
  
right_label <- sum1 %>%
  filter(gender=='Male')

left_label <- sum1 %>% 
  filter(gender=="Female")

sum1 %>% 
ggplot(aes(x=mob_final,
             y=prop,
             colour=gender))+
  geom_point(aes(size=0.3))+
  geom_errorbar(aes(ymin=conf_up,
                    ymax=conf_low),
                width=0.3,
                size=0.3)+
  labs(subtitle= "Estimates contain 95% confidence intervals (Standard error*1.96)",
       x="Mobility type",
       y="Proportion of respondents",
       caption="RLMS rounds 2011 to 2015 \nAuthors own calculation")+
  guides(size=FALSE)+
  geom_text(data = right_label, aes(color = gender, label = round(prop, 3)),
                               size = 3, hjust = -.6) +
  geom_text(data = left_label, aes(color = gender, label = round(prop, 3)),
            size = 3, hjust = 1.8)+
  theme_minimal()

```


Figure 2 considers the same information, but splits the date by age of respondent. 
Younger respondents are more more mobile and more likely to list both promotion and firm exit than older groups. 
Both of these measures gradually decline as we consider older groups. 
Noteably, the gender difference in firm exits dissapears for older groups, but remains for the youngest group. 


```{r, fig.cap="Age and gender differences in mobility"}

sum2 <- ru3 %>%
filter(!is.na(mob_final)) %>% 
  group_by(gender, age_groups , mob_final) %>%
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n),
         prop_dev =prop*(1-prop),
         se=sqrt(prop_dev/n),
         conf_up = prop + (se*1.96),
         conf_low = prop -(se*1.96))  
  
right_label <- sum1 %>%
  filter(gender=='Male')

left_label <- sum1 %>% 
  filter(gender=="Female")

sum2 %>% 
ggplot(aes(x=mob_final,
             y=prop,
             colour=gender))+
  geom_point(aes(size=0.3))+
  geom_errorbar(aes(ymin=conf_up,
                    ymax=conf_low),
                size=0.3)+
  labs(subtitle= "Estimates contain 95% confidence intervals (Standard error*1.96)",
       x="Mobility type",
       y="Proportion of respondents",
       caption="RLMS rounds 2011 to 2015 \nAuthors own calculation")+
  guides(size=FALSE)+
  facet_wrap(~age_groups)+
  scale_x_discrete(labels = abbreviate)+
  theme_minimal()

```

Both firgures make an important point, there are gender and age effects tied job mobility, with younger workers being more mobile than older workers, and younger men being more mobile than younger women. 
We now turn to the wage differences between these groups. 
As mentioned throughout, we will not consider the effect of promotion or exit on wages. Instead we will look at group differences in wages, by gender. 

```{r, fig.cap="Wage differences between gender and mobility groups"}
sum3 <- ru3 %>%
  filter(!is.na(mob_final),
         !is.na(wage)) %>% 
  group_by(gender, mob_final) %>% 
  summarise(wage_mean = mean(wage, na.rm=TRUE),
            wage_sd = sd(wage, na.rm = TRUE),
            n= n(),
            wage_se = wage_sd/sqrt(n)) 
 

sum3 %>% 
ggplot(aes(x=mob_final, 
             y= wage_mean,
             col=gender))+
  geom_line(aes(group=gender))+
  geom_point(aes(group=mob_final,
                 size=3))+
  geom_errorbar(aes(ymin=wage_mean-(wage_se*1.96),
                    ymax=wage_mean+(wage_se*1.96)),
                width=0.25)+
  scale_y_continuous(labels = scales::comma)+
  labs(subtitle= "Estimates contain 95% confidence intervals (Standard error*1.96)",
       y= "Average monthly wages in rubles",
       x="Mobility type",
       caption="RLMS rounds 2011 to 2015 \nAuthors own calculation")+
  guides(size=FALSE)+
  geom_text(data = sum3, aes(color = gender, label = round(wage_mean, 0)),
            size = 3, hjust = -.6)+
  theme_minimal()


```

Figure 3 captures three important differences. 
First, there is a gender difference tied to pay, with men earning more than women in each mobility cateogry. 
Second, there is difference in pay between mobility groups, with respondents who experience a promotion in the survey year citing significantly higher wages than those with the same job with the same employer.
Third, there is a pay penalty among respondents who exit a firm, but this penalty is only apparent for men. 
This effect likely stems from women's pay reaching a "floor".
We consider the age differences in this effect below.


```{r, fig.cap="Wage differences between age, gender, and mobility groups"}
sum4 <- ru3 %>%
  filter(!is.na(mob_final),
         !is.na(wage)) %>% 
  group_by(gender, age_groups, mob_final) %>% 
  summarise(wage_mean = mean(wage, na.rm=TRUE),
            wage_sd = sd(wage, na.rm = TRUE),
            n= n(),
            wage_se = wage_sd/sqrt(n)) 
 

sum4 %>% 
ggplot(aes(x=mob_final, 
             y= wage_mean,
             col=gender))+
  geom_line(aes(group=gender))+
  geom_point(aes(group=mob_final,
                 size=3))+
  geom_errorbar(aes(ymin=wage_mean-(wage_se*1.96),
                    ymax=wage_mean+(wage_se*1.96)),
                width=0.25)+
  scale_y_continuous(labels = scales::comma)+
  labs(subtitle= "Estimates contain 95% confidence intervals (Standard error*1.96)",
       y= "Average monthly wages in rubles",
       x="Mobility type",
       caption="RLMS rounds 2015 to 2011 \nAuthors own calculation")+
  guides(size=FALSE)+
  facet_wrap(~age_groups)+
  scale_x_discrete(labels = abbreviate)+
  theme_minimal()


```

Thinking of differences between age groups in Figure 4, respondents aged 36-55 appear to gain the most from promotion, although the premium only appears to apply to men. Surprisingly, older groups report the lowest income and also the lowest return on mobilty, even though the chosen outcome relies on wages, and not income more broadly. Older women appear to make significant gainst when citing promotion, however the large confidence intervals suggest this premium is not significantly different from other age groups.

Generally we find a difference between age and gender groups in the likelihood of mobility.
We also find that respondents who cite mobility tend to have higher wages in the years when they experience mobility. 
However, the graphs above only lay out the baseline differences between men and women, without considering gender differences in hours worked, experience, and occupations. 
The section below turns to these explanations, predicting mobility events using the controls described throughout.  

## Multinomial logistic regression

We now consider some explanatory measures tied tojob mobility differences. 


```{r, warning=F, echo=FALSE, message=FALSE, error=FALSE, tidy=TRUE}
d1<- ru3 %>% 
  filter(age_groups=="0-35")
d2<- ru3 %>% 
  filter(age_groups=="36-55")


m1 <- multinom(mob_final ~  gender + age  + tenure + super + factor(esec_simple) +factor(marr) + hours, data=d1)
m2 <- multinom(mob_final ~  gender + age  + tenure + super + factor(esec_simple) +factor(marr) + hours, data=d2)

t1<- tidy(m1) %>% 
  select(-statistic, -y.level) %>% 
  mutate(signif = stars.pval(p.value))

t2 <- tidy(m2) %>% 
  select(-y.level, -term, -statistic) %>% 
  mutate(signif = stars.pval(p.value)) 


kable(cbind(t1, t2), digits=3,
      booktabs=TRUE) %>% 
  kable_styling(latex_options = c("scale_down","repeat_header")) %>% 
  add_header_above(c(" " = 1, "Age 18-35" = 4, "Age 36-55" = 4)) %>% 
  pack_rows("Promotion", 1, 10) %>% 
  pack_rows("Lateral change", 11, 20) %>% 
  pack_rows("Exit", 21, 30) %>% 
  footnote(number = c("RLMS rounds 20-25. Models consider two separate age groups", "*** p < 0.001, ** p < 0.01, * p < 0.05, . p < 0.1"))



```


We discuss the main findings in turn, focusing on one mobility outcome at a time. 
Regarding promotions, we first look at models for respondents aged 18-35. 
There is a significant gender between young men and women, with women reporting lower odds (0.812) of experiecing a promotion when compared to men (1.0).
This difference remains even when we control for differences in age, firm tenure, supervisory position, the occupation of respondents, and their marital status.
Women are less likely to cite a promotions when compared to men, even when considering the differences above. 
Beyond this, the controls have their own imapct on promotions. 
Younger respondents are more likely to receive a promotion than older respondents, as are respondents with less firm experience comapred to respondents with more firm experience. 
Respondents in supervisor positions (1.0) are more likely to cite a promotion when comapred to respondents without a superisor position (0.225). 
Finally, regarding occupations, those at the lower end of the occuaptional hierarchy cite lower odds of experiencing a promotion, when compared to the upper half of the occupational distribution.
In general, promotions occur most often among younger males with supervisory positions, in high occupations. 
Thus the early stages of one's career are crucial to establishing chance of promotion, but even here, women are less likely to receive these when comapred to men. 

We now turn to respondents aged 36-55. 
The previous gender difference no longer applies to these respondents.
Women are still less likely to cite a promotion (0.092) when comapred to men (1.0), but this difference is no longer statistically significant. 
However, the remaining controls have an impact on promotion. 
Younger respondents are again most likely to cite promotion. 
As are respondents with less tenure than more established workers. 
Those who are in supervisor positions (1.0) are more likely to experience promotion when compared to respondents without a supervisory position (0.246).
There are minor differences between occuaptions in terms of promotion, but generally those in lower occupations are unlikely to cite promotion, when compared to higher occupational categories. 

Regarding lateral mobility within the firm, we again consider younger respondents first. 
We find no gender difference between men and women regarding their odds of experiencing lateral change.
There is a minor age effects $(p < 0.1)$, which suggests that older respondents are less likely to change jobs laterally, when compared to older respondents.
However, this effect is minor. 
Respondents with higher firm tenure are also less likely to experience voluntary mobility when comapred to newer respondents.
There are no other significant effects in the model for younger respondents.

Considering workers aged 36-55, the only significant estimates are the many occupational differences in lateral moves. 
These suggest that both high and low occupational categories have higher odds of changing jobs within the firm when comapred to respondents at the top of the occupational categories (professionals).
There is a second curious result which suggests that separated respondents and two times more likely to move laterally when comapred to respondents who never married. 
This could stem from the need to work more flexible hours within the firm, and likely imapcts women more than men. 
This is not the primary goal of this article. 

Finally, we consider respondents who cite an employment exit compared to those who remain in the same job with the same employer. 
Starting with younger respondents (aged 18-35), we find a gender difference in firm exits. 
Women are less likely to cite an exit (0.715) comapred to men (1.0).
This effect remains even when we control for age, firm level tenure, supervisory positions, occupational differences, and marital status. 
Further, the controls themselves have their own impact on job exit. 
the models suggest that older respondents (those closer to 35) are more likely to exit a firm when comapred to younger respondents.
However firm tenure has a strong and negative effect on exit, those who remain with a firm for longer and less likely to cite a firm exit.
The remaining measures have no effect on predicting firm exit. 

Considering respondents who are aged 36-55, we find a simialr pattern. Once again women are less likely to cite firm exit when compared to men, even when controlling for a range of factors associated with firm level exit. 
One major difference between these respondents and younger respondents is that the effect of age is not the opposite.
Among these respondents it is older respondents who are less likely to exit, when compared to younger respondents. 
The effects of firm level tenure, however, is the same. 
As workers spend more time in the firm, they are less likely to leave, while respondents with less tenure are more likely to leave. 
The remaining measures are insignificant. 

We note two important findings from the models above. First, mobility is most prominent and most common in the earliest stages of one's career. 
This suggests that once respondents secure their positions, they likely hold on to them moving only rarely after the initial period of adjustment.
Further within the firm, young respondents with less tenure are more likely to cite a promotion, but those with higher tenure becomegraudually less likely to exit the firm, suggesting limited mobility.
Second, young women are less likely to  experience a promotion when compared to young men, and young women are less likely to experience a firm exit when comapred to young men. 
This remains even when we controls for a range of gender differences between men and women. 
This finding contradicts a number of previous authors, both those writing about Russia and those writing about other countries.
Summarise these...


## Fixed effects estimation

Having established the major age and gender differences between quits and promotions, we now turn to their effects on earnings. 
As mentioned we consider linear fixed-effects estimation throughout.
We start by considering respondents aged between 18-35, before considering respondents aged 36-55.
We will see two different results for younger and older workers, but for now we focus on the gender differences within age groups.


```{r}
df_m<- ru3 %>% 
 mutate(ln_wage = log(wage),
       t= year - 2011) %>% 
filter(gender=="Male",
        ln_wage !=-Inf,
        age_groups=="0-35")



df_m <- pdata.frame(df_m, index=c("idind", "round"))

pm_m <- plm(ln_wage ~ 1 + mob_final + age + tenure + super + factor(esec_simple) +factor(marr) + hours, 
                   data = df_m,
                   index = c("idind", "round"), 
                    model = "within")




df_f<- ru3 %>% 
 mutate(ln_wage = log(wage),
       t= year - 2011) %>% 
filter(gender=="Female",
        ln_wage !=-Inf,
        age_groups=="0-35")




df_f <- pdata.frame(df_f, index=c("idind", "round"))

pm_f <- plm(ln_wage ~ 1 +  mob_final + age + tenure + super + factor(esec_simple) +factor(marr) + hours, 
                   data = df_f,
                   index = c("idind", "round"), 
                    model = "within")


## combine male and female results

male_results <- tidy(pm_m) %>% 
  select(-statistic) %>% 
  mutate(signif = stars.pval(p.value))

female_results <- tidy(pm_f) %>% 
  select(-term, -statistic) %>% 
  mutate(signif = stars.pval(p.value)) 


##knit

kable(cbind(male_results, female_results), digits=3,
      booktabs=TRUE) %>% 
  kable_styling(latex_options = c("scale_down","repeat_header")) %>% 
  add_header_above(c(" " = 1, "Male Results (aged 18-35)" = 4, "Female results (aged 18-35)" = 4)) %>% 
  footnote(number = c("RLMS rounds 20-25. Models consider gender separately", "*** p < 0.001, ** p < 0.01, * p < 0.05, . p < 0.1"))



```

The table above focuses mainly on respondents aged 18-35. 
We first consider the effects of job mobility on men.
Respondents who experience a promotion report no change in monthly wages. 
Although the estimate is positive, it is not significant. 
Further, lateral changes, and firm exits have a negative effect on monthly wages, suggesting that workers who experience these changes see a fall in earnings, at least in the short term. 
Male respondents see a strong positive effect of ageing on earnings, but an insignificant effect of job tenure on earnings. 
There is a relationship between moving from a supervisory role to a non-supervisory role, suggesting that respondents who can become supervisors see a positive change in earnings after the change. 
Controlling for occupational change has not effect on earnings, which is a strange result, but reflects the fact that occupational mobility is uncommon among younger workers. 
Finally there is a weak but positive effect $(p>0.1)$ of marriage on earnings.
This suggests that respondents who move from being unmarried to married are able to secure a minor premium after the change. 
Such a result is common in sociological studies of the labour market. 

We now consider the resutls for female respondnets which are quite different from that of male respondents. 
Women see significant premiums tied to every type of mobility. 
For example, women who receive a promotion typically see a 17 per cent premium in their monthly earnings. 
In a similar vein women who move lateraly within the firm see a 28 per cent increase in pay after the change.
Lastly, women who exit the firm for a new employer report an 8 per cent rise in pay.
These findings sugegst tha mobility isparticualrly important for young women, who in the early stages of their career must quickly establish themselves by changing positions in a firm, or between firms. 
The same results do not emerge for men.
Beyond mobility, the controls have significant effects on monthly wages. 
First, age has a positive and significant effect on women's earnings, similar to the effect in men's models. 
However, tenure appears to have a negative and significant effect on earnings.
This suggests that mobility is particularly important in the early stages of one's career, as staying in the same position limits the sharp increase in earnings that are tied to age for young women.
As with men, there is a pay penalty tied to moving to a position without being a supervisor, which suggests a pay premium to becoming a supervisor. 
There are also significant premiums and penalties tied to occupational mobility. 
Stangely, the marriage premium that we noted for men, is not a penalty for young women. 
This penalty is also statistically significant. 
Separation also appears to have a minor but negative effect $(p<0.1)$. 







```{r}
df_m<- ru3 %>% 
 mutate(ln_wage = log(wage),
       t= year - 2011) %>% 
filter(gender=="Male",
        ln_wage !=-Inf,
        age_groups=="36-55")



df_m <- pdata.frame(df_m, index=c("idind", "round"))

pm_m <- plm(ln_wage ~ mob_final + age + tenure + super + factor(esec_simple) +factor(marr) + hours, 
                   data = df_m,
                   index = c("idind", "round"), 
                    model = "within")




df_f<- ru3 %>% 
 mutate(ln_wage = log(wage),
       t= year - 2011) %>% 
filter(gender=="Female",
        ln_wage !=-Inf,
        age_groups=="36-55")




df_f <- pdata.frame(df_f, index=c("idind", "round"))

pm_f <- plm(ln_wage ~ mob_final + age + tenure + super + factor(esec_simple) +factor(marr) + hours, 
                   data = df_f,
                   index = c("idind", "round"), 
                    model = "within")


## combine male and female results

male_results <- tidy(pm_m) %>% 
  select(-statistic) %>% 
  mutate(signif = stars.pval(p.value))

female_results <- tidy(pm_f) %>% 
  select(-term, -statistic) %>% 
  mutate(signif = stars.pval(p.value)) 


##knit

kable(cbind(male_results, female_results), digits=3,
      booktabs=TRUE) %>% 
  kable_styling(latex_options = c("scale_down","repeat_header")) %>% 
  add_header_above(c(" " = 1, "Male Results (aged 36-55)" = 4, "Female results (aged 36-55)" = 4)) %>% 
  footnote(number = c("RLMS rounds 20-25. Models consider gender separately", "*** p < 0.001, ** p < 0.01, * p < 0.05, . p < 0.1"))



```

some more thoughts and feelings 


### scraps

 this occured because workers were not being paid, rather than their desire to improve pay between two employers [@clarke2002market]. In fact Clarke [-@clarke2002market] noted a sharp rise in Russia's real wages during reforms, that also lead to significant variance in wages. He noted however, that the beneficiaries of these wages were respondents who remained in relatively successful firms which protected workers from instability. 



\newpage
# References


