---
title: "Gender Differences in Russia's Job Mobility and Its Rewards"
author: "Author"
date: "02 December 2019"
output: pdf_document 
bibliography: 'C:/Users/Ivan.Privalko/Desktop/R projects/ru_mobility/biblio/ru1.bib'
toc: TRUE
---


```{r, echo=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(forcats) 
library(gtools)
library(knitr)
library(kableExtra)
library(nnet)
library(broom)
library(plm)


knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)

load("C:/Users/Ivan.Privalko/Desktop/R projects/ru_mobility/genderpaper.RData")

ru3 <- ru3 %>% 
  mutate(age_groups=case_when(
    age %in% 0:35 ~  "0-35",
    age %in% 36:55 ~ "36-55",
    age %in% 56:100 ~ "56+"
  ))


```
\newpage
# Abstract

This article uses data from Russia's Longitudinal Monitoring Survey to consider age and gender differences in the probability and consequences of quits and promotions.
Russia's relatively liberalised labour market should have high rates of job mobility.
However,  IMF figures suggest that mobility in the early 2010's is on par with mobility recorded in the 1980's.
Beyond this, little has been said on who is mobile in Russia and whether mobility has any impact on wages, once personal characteristics of movers are omitted.
Results show significant difference between age and gender groups regarding both exits and internal promotions.
Regarding promotions, when several personal and job characteristics are held constant, instances of promotion are higher for men when compared to women. 
Further, age is a strong predictor of the likelihood of promotion, even among 18-35 year olds, promotions are less likely for older respondents (closer to 35) than they are for younger respondents (closer to 18).
Regarding firm exits, when several personal and job characteristics are held constant, instances of exit are again higher for men when compared to women. 
Here too, age differences suggest that younger respondents are more likely to experience exit when comapred to older respondents, even within groups of respondents aged 18-35.
In general, the early stages of a respondent's career are marked by periods of high mobility in Russia, which is similar to the experience of young workers in other countries.
However, after this period, mobility becomes increasingly unlikely.
Part of this result could stem from the premiums tied to promotion. 
I find that after controlling for individual heterogeneity (through the use of fixed effects estimation), firm exits do not improve immediate wages although promotions contain significant wage premiums, at least in the short term. 
Results help to understand processes of inequality in wages and conditions that occur due to sorting, and the importance of promotions (internal job changes with the same employer) as "life chances" which improve earnings in the immediate sense.
Gender differences in securing these life chances may help to understand wider gender gaps in earnings, which emerge later. 

# Introduction 

Russia's liberalised economy gave workers more employment options and a wider spectrum of potential wages and conditions [@imf1991study; @clarke2002market]. 
As a result, theorists assumed that workers would increase their job mobility in order to correct their earnings and conditions [@keith1995wage; @keith1997job; @cha2014job; @clarke2002market].
Such corrections are an important part of *job searching* and *job matching* theories of inequality [@schmelzer2012consequences; @schmelzer2011consequences].
Despite these predictions, Russian job mobility rates in 2010 were lower than they were in the 1980's under the Soviet Union [@imf1991study]. 
Although the labour market saw significant churn immediately after reforms in the 1990's, this rate gradually declined and returned to previous levels, with workers preferring to remain in the same job with the same employer [@clarke2000closure]. As a result, Russia presents an itneresting puzzle to researchers, one where wage inequality is relatively high, but job mobility is relatively low.

This article has two aims. 
First, we will focus on predicting mobility, estimating group differences in who moves and who remains in the same job with the same employer. 
Do gender differences in Russia's job mobility exist, and if they do are these the product of gender, or gender differences in occupational sorting or firm size? 
Further, is mobility confined only to younger workers, or do certain occupations contain mobility patterns that cannot be explained by age differences between workers alone?
Second, having established these differences, we will consider the effects that moving has on wages. 
Does mobility reward workers, in that after we consider the characteristics of those most likely to move, do premiums exist in the earnings of workers, and do those differ by mobility within and between firms?
Again, I pay attention to the gender differences in the effects of mobility on earnings, asking whether men and women benefit equally from the change.

We will use the Russia Longitudinal Monitoring Survey (RLMS) throughout  as a representative sample of the country's workforce. 
This approach has two strengths which are relevant to the wider discussion of job mobility. 
First, the survey explicitly measures firm exits and promotions which occurred in the last twelve months.
Previous authors discussing mobility often operationalise promotions and exits in complex ways [@kalleberg2001satisfied; @keith1995wage; @keith1997job].
Second, by using longitudinal data we will focus on the effects of mobility on wages, instead of comparing the wages of movers to non-movers [@cha2014job; @reichelt2017occupational].

The article has three main findings .... summarise here.

The article is structured as follows. The first section summarises the theoretical literature on job mobility. The second section summarises the empirical findings between age, gender, and job mobility. This section also presents Russia as an important case in the studies of job mobility. The third section summarises the data and the approach of estimating the likelihood of mobility, and the effect of mobility. Section four presents the results, while a brief discussion concludes the article. 

# Review

## Job searching and Job matching 

Studies of job mobility rely on two theoretical approaches, *job-searching* and  *job-matching*. 
Both suggest that poor work environments either from inadequate pay or inadequate conditions, push workers to search for new positions [@kalleberg2001satisfied; @gesthuizen2008mismatching; @schmelzer2011consequences]. 
The *job searching* approach considers the *Reservation Wage*. 
This is the minimum wage a worker accepts before moving to a new position, given their skill set. 
*Reservation wage X* draws employees out of unemployment and is the minimum wage that he or she will accept before taking a job.
*Reservation wage Y* where (X < Y) draws employees to new positions either within or between firms; it is the minimum wage that a worker will accept before changing positions.
If a previously unemployed worker accepts *Wage X* that is below *Wage Y*, they are assumed to continue searching until they secure *Wage Y* [@schmelzer2010consequences; @burdett1978theory].
*Wage Y* is typically greater than *Wage X* for several reasons including the fact that job-searching is costly to workers. 
Most importantly however, workers searching for new positions have access to wider networks of information and wider knowledge about possible reservation wages. 
There may be gender differences in these networks, which can lead to differences in reservation wages after the change.

On the other hand, the *job-matching* approach is one where workers take time to evaluate whether they have a good amtch between their skills and their tasks at work. 
Since work is an experience good, workers must try several jobs before finding a match.
The large rates of mobility that workers see in the early stages of their career, is them navigating the labour market, trying individual employers and jobs and comparing their experience with rpevious experiences [@sorensen1977structure; @thurow1975generating]. 

How do the theories above explains the large differences in mobility between workers? 
Both understand that workers are most mobile in the early stages of their career. 
However, in the job searching approach, individuals try to secure their reservation wages as they establish themselves in the labour market and gain the experience and networks needed to evaluate new positions. 
As they gradually close the gap between their given wage and their reservation wage, they become less mobile [@sorensen1977structure; @kalleberg1979sociology; @sorensen1975structure].
The job-matching approach is slightly different in that it does not expect a wage premium tied to mobility, only that the subjectively evaluated match between a worker and her position should improve following mobility. 
This approach also anticipates the large rate of mobility in the early stages of a person's career, but in this mechanism workers are "experiencing" the type of work they would like to do. 
It is possible that this mechanism affects men and women differently, in that men would be more driven by wages, and women be more driven by work life balance or their experience at work. This is likely especially true if women hold the majority of care responsibilities in the home [@keith1997job].

## Gender differences in mobility and its returns

Considering both theories together, neither explicitly notes a gender difference in job mobility. 
Despite this, authors routinely find this difference  [@blau1981race; @keith1995wage; @keith1997job; @keith1999returns]. 
Blau and Duncan [-@blau1981race] note young women are more likely to quit when compared to men. 
Keith and McWilliam [-@keith1995wage] note that economic quits happen at a similar rate for men and women, but that family related changes are more common among young women, when compared to young men. 
They also find that men are more likely to experience job loss, when compared to young women, suggesting that gender differences stem mostly from different types of mobility, rather than a gender difference in wage returns from similar mobility types [@keith1995wage; @keith1999returns].
In short, part of the voluntary job mobility carried out by women is closer tied to job matching than job searching, in this way Keith and McWilliams [-@keith1995wage; -@keith1999returns] suggest women find positions where they are better able to combine family roles and work. 
Gesthuizen finds some evidence to support this, reporting that dissatisfaction with the job match and dissatisfaction with hours increases women's chances of voluntary mobility to a new employer.
This effect does not emerge for men. 
Further, subjective dissatisfaction with wages also pushes women to inter-firm mobility, but not men, which would support the job searching hypothesis. 
In discussing the result Gesthuizen [] notes that mobility may be “less costly” for women, in that men are more likely typically “breadwinner” roles, with less room
Keith and McWilliams [-@keith1999returns] also note a gender difference in job searching behaviour, reporting men are more likely to explore new positions than women. Job searching behaviour was particularly important as it brought a significant wage premiums for both men and women who engaged in job mobility. 

Beyond gender differences in mobility types, mobility may have different effects on men and women. 
Keith and McWilliams [-@keith1999returns, -@keith1995wage; -@keith1997job] find no gender difference in the effects of job mobility on wages, assuming the mobility occurs for economic reasons. 
Further, they find that men and women benefit roughly equally from the interaction between job search and job mobility.
Gesthuizen [] also finds no significant difference between men and women in the premium tied to mobility.
Cha [-@cha2014job] notes a significant difference in the effect of mobility on earnings between women with and without children, but one that runs opposite to the described pattern. 
Women with children see few rewards to mobility when compared to women without children, who experienced significant premiums in earnings.
The study doesn't consider men.
Kronberg [-@kronberg2013stay] finds this specific difference, reporting that men gain more from voluntary mobility to a new employer when compared to women, although see a premium also. 


Although other authors find no gender differences between men and women in the wage effects of "economic quits" between young men and women in the early stages of their career [@keith1995wage]. This suggests that the wage or conditions premium described above applies more often to men than to women, even when removing gender differences in types of mobility. 
Some authors argue that the lack of a gender  


## Russia's labour market, as a case

Russia's fast transition to a post-soviet economy had positive and negative effects on gender differences in its labour market. 
Although transition increased women's participation and employment, it also increased levels variation in women's job quality [@gerber2006dynamic]. 
Given the pace of market transition, this increased variation did not stem from a decline in human capital, but the *"introduction of market institutions"* [@gerber2006dynamic].
Since the liberalization of the labour market, authors have argued 
Despite the theories above, Russian levels of job mobility actually fell in the 1990's, to levels below that of 1980's soviet job mobility, at least in measures of inter-firm mobility (voluntary quits to new employers). 


# Methodology

This article uses five rounds of Russia's Longitudinal Monitoring Survey (Rounds 20-24), a representative and longitudinal dataset of Russia's population [@kozyreva2016data]. We focus only on those who are in employment and 

# Results

This section is split into three parts. 
First, I present gender differences in mobility, and the wages tied to certain mobility responses.
Second, I explore gender differences in mobility using multinomial logistic regression.
Finally, we explore the effects of mobility on earnings using fixed-effects linear regression.

## Descriptive statistics

Figure 1  considers gender differences in mobility types focusing on observations rather than individuals. 
Overall, respondents are largely immobile although women (0.814) are less mobile compared to men (0.762). 
Both men and women (0.044) are equally likely to list a promotion in a given year, although promotions are uncommon in general. 
Respondents are much more likely to list exiting a firm, with men (0.182) listing more exits than women (0.130).
Lateral changes within the firm are the least common and do not appear to have a gender difference. 
Considering the confidence intervals in Figure 1, it seems gender differences are most prominent in terms of firm exits.
Unfortunately we are not able to discern voluntary from involutnary exits, which are particualrly important for discussions of reservations wages.
However, this category, along with promotions, likely contains the job mobility which is driven by reservation wages.
 

```{r, fig.cap="Proportion of respondents citing mobility type by gender"}

sum1 <- ru3 %>%
filter(!is.na(mob_final)) %>% 
  group_by(gender, mob_final) %>%
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n),
         prop_dev =prop*(1-prop),
         se=sqrt(prop_dev/n),
         conf_up = prop + (se*1.96),
         conf_low = prop -(se*1.96))  
  
right_label <- sum1 %>%
  filter(gender=='Male')

left_label <- sum1 %>% 
  filter(gender=="Female")

sum1 %>% 
ggplot(aes(x=mob_final,
             y=prop,
             colour=gender))+
  geom_point(aes(size=0.3))+
  geom_errorbar(aes(ymin=conf_up,
                    ymax=conf_low),
                width=0.3,
                size=0.3)+
  labs(subtitle= "Estimates contain 95% confidence intervals (Standard error*1.96)",
       x="Mobility type",
       y="Proportion of respondents",
       caption="RLMS rounds 2011 to 2015 \nAuthors own calculation")+
  guides(size=FALSE)+
  geom_text(data = right_label, aes(color = gender, label = round(prop, 3)),
                               size = 3, hjust = -.6) +
  geom_text(data = left_label, aes(color = gender, label = round(prop, 3)),
            size = 3, hjust = 1.8)

```


Figure 2 considers the age differences tied to the changes in Figure 1. We note that younger respondents are more likely to experience promotions and exits. Both of these measures gradually decline as we consider older groups. Noteably, gender differences in quits dissapear for older respondents, but remain for younger respondents. This effect stems from the sharp decline in men's likelihood of quitting. 

```{r, fig.cap="Age and gender differences in mobility"}

sum2 <- ru3 %>%
filter(!is.na(mob_final)) %>% 
  group_by(gender, age_groups , mob_final) %>%
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n),
         prop_dev =prop*(1-prop),
         se=sqrt(prop_dev/n),
         conf_up = prop + (se*1.96),
         conf_low = prop -(se*1.96))  
  
right_label <- sum1 %>%
  filter(gender=='Male')

left_label <- sum1 %>% 
  filter(gender=="Female")

sum2 %>% 
ggplot(aes(x=mob_final,
             y=prop,
             colour=gender))+
  geom_point(aes(size=0.3))+
  geom_errorbar(aes(ymin=conf_up,
                    ymax=conf_low),
                size=0.3)+
  labs(subtitle= "Estimates contain 95% confidence intervals (Standard error*1.96)",
       x="Mobility type",
       y="Proportion of respondents",
       caption="RLMS rounds 2011 to 2015 \nAuthors own calculation")+
  guides(size=FALSE)+
  facet_wrap(~age_groups)+
  scale_x_discrete(labels = abbreviate)

```

Generally, there is a gender and age effect tied job mobility, with younger workers being more mobile than older workers, and younger men being more mobile than younger women. We now turn to the wage differences between these groups. Importantly, we will not consider the effect of promotion or exit on wages, focusing instead on the average wage associated with a given gender and a given mobility type. 



```{r, fig.cap="Wage differences between gender and mobility groups"}
sum3 <- ru3 %>%
  filter(!is.na(mob_final),
         !is.na(wage)) %>% 
  group_by(gender, mob_final) %>% 
  summarise(wage_mean = mean(wage, na.rm=TRUE),
            wage_sd = sd(wage, na.rm = TRUE),
            n= n(),
            wage_se = wage_sd/sqrt(n)) 
 

sum3 %>% 
ggplot(aes(x=mob_final, 
             y= wage_mean,
             col=gender))+
  geom_line(aes(group=gender))+
  geom_point(aes(group=mob_final,
                 size=3))+
  geom_errorbar(aes(ymin=wage_mean-(wage_se*1.96),
                    ymax=wage_mean+(wage_se*1.96)),
                width=0.25)+
  scale_y_continuous(labels = scales::comma)+
  labs(subtitle= "Estimates contain 95% confidence intervals (Standard error*1.96)",
       y= "Average monthly wages in rubles",
       x="Mobility type",
       caption="RLMS rounds 2011 to 2015 \nAuthors own calculation")+
  guides(size=FALSE)+
  geom_text(data = sum3, aes(color = gender, label = round(wage_mean, 0)),
            size = 3, hjust = -.6)


```

Figure 4 captures thre eimportant differences. First, there is a gender difference tied to pay, with men earning more than women in each mobility cateogry. Second, there is a bility difference in pay, with respondents who experience a promotion in the survey year citing significantly more in terms of average earnings, compared to respondents citing a different mobility type in a given survey year. Third, there is a pay penatly among respondents who cite employment exit, but this difference is only significant for men. This effect likely stems from women's pay reaching a "floor". We consider the age differences in this effect below.


```{r, fig.cap="Wage differences between age, gender, and mobility groups"}
sum4 <- ru3 %>%
  filter(!is.na(mob_final),
         !is.na(wage)) %>% 
  group_by(gender, age_groups, mob_final) %>% 
  summarise(wage_mean = mean(wage, na.rm=TRUE),
            wage_sd = sd(wage, na.rm = TRUE),
            n= n(),
            wage_se = wage_sd/sqrt(n)) 
 

sum4 %>% 
ggplot(aes(x=mob_final, 
             y= wage_mean,
             col=gender))+
  geom_line(aes(group=gender))+
  geom_point(aes(group=mob_final,
                 size=3))+
  geom_errorbar(aes(ymin=wage_mean-(wage_se*1.96),
                    ymax=wage_mean+(wage_se*1.96)),
                width=0.25)+
  scale_y_continuous(labels = scales::comma)+
  labs(subtitle= "Estimates contain 95% confidence intervals (Standard error*1.96)",
       y= "Average monthly wages in rubles",
       x="Mobility type",
       caption="RLMS rounds 2015 to 2011 \nAuthors own calculation")+
  guides(size=FALSE)+
  facet_wrap(~age_groups)+
  scale_x_discrete(labels = abbreviate)


```

Thinking of the patterns in Figure 4, respondents appear to gain the most from mobility aged 36-55, although the premium only appears to apply to men. Surprisingly, older groups report the lowest income and also the lowest return on mobilty. Older women appear to make significant gainst when citing promotion, however the large confidence intervals suggest this premium is not significantly different from other age groups.

## Multinomial logistic regression

We now consider some explanatory measures tied tojob mobility differences. 


```{r, echo=FALSE, warning=FALSE}
d1<- ru3 %>% 
  filter(age_groups=="0-35")
d2<- ru3 %>% 
  filter(age_groups=="36-55")

m1 <- multinom(mob_final ~  gender + age + tenure + super + factor(isco08), data=d1)
m2 <- multinom(mob_final ~  gender + age + tenure + super + factor(isco08), data=d2)


t1<- tidy(m1) %>% 
  select(-statistic, -y.level) %>% 
  mutate(signif = stars.pval(p.value))

t2 <- tidy(m2) %>% 
  select(-y.level, -term, -statistic) %>% 
  mutate(signif = stars.pval(p.value)) 




kable(cbind(t1, t2), digits=3,
      booktabs=TRUE) %>% 
  kable_styling(latex_options = c("scale_down","repeat_header")) %>% 
  add_header_above(c(" " = 1, "Age 18-35" = 4, "Age 36-55" = 4)) %>% 
  pack_rows("Promotion", 1, 4) %>% 
  pack_rows("Lateral change", 5, 8) %>% 
  pack_rows("Exit", 9, 12) %>% 
  footnote(number = c("RLMS rounds 20-25. Model considers all respondents together", "*** p < 0.001, ** p < 0.01, * p < 0.05, . p < 0.1"))






```


## Fixed effects estimation

We now turn to the quetion of how mobility affects earnings
```{r}
#df_m<- ru3 %>% 
 # mutate(ln_wage = log(wage),
  #       t= year - 2011) %>% 
  #filter(gender=="Male",
 #        ln_wage !=-Inf) 



#df_m <- pdata.frame(df_m, index=c("idind", "round"))
#df_m


#df_f<- ru3 %>% 
 # mutate(ln_wage = log(wage),
  #       t= year - 2011) %>% 
  #filter(gender=="Female",
   #      ln_wage !=-Inf,
#         age_groups=="36-55") 

#ru3 %>% 
#  count(esec_simple) 

#pm_fe <- plm(ln_wage ~ mob_final + age + tenure +super + firm_size + factor(esec_simple) +  t, 
 #                   data = df_m,
  #                  index = c("idind", "round"), 
   #                 model = "within")


#summary(pm_fe)

#tidy(pm_fe) %>% 
#mutate(sig=stars.pval(p.value)) %>% 
#kable(digits = 3)


```


### scraps

 this occured because workers were not being paid, rather than their desire to improve pay between two employers [@clarke2002market]. In fact Clarke [-@clarke2002market] noted a sharp rise in Russia's real wages during reforms, that also lead to significant variance in wages. He noted however, that the beneficiaries of these wages were respondents who remained in relatively successful firms which protected workers from instability. 



# References


