---
title: "Age and Gender Difference in Quitting and Its Rewards"
author: "Ivan Privalko"
date: "02/12/2019"
output: pdf_document
bibliography: 'C:/Users/Ivan/Dropbox/Biblio/bibblio.bib'
---


```{r, echo=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(forcats) 
library(knitr)
library(kableExtra)
library(nnet)
library(broom)


knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)

load("C:/Users/Ivan/Desktop/dir/papers/ru_mobility/genderpaper.RData")

ru2 <- ru2 %>% 
  mutate(age_groups=case_when(
    age %in% 0:35 ~  "0-35",
    age %in% 36:55 ~ "36-55",
    age %in% 56:100 ~ "56+"
  ))

```

# Abstract

The article uses data from Russia's Longitudinal Monitoring Survey to estimate age and gender differences in the probability and consequences of job mobility. 
uit rates are similar for young men and women, but finds gender differences among older workers. When several personal and job characteristics are held constant, however, the quit rates of young men and women are about the same, suggesting gender differences can be explained by early career sorting. When controlling for individual heterogeneity (through the use of fixed effects estimation), it is found that, for all race and sex groups, quitting does not improve current wages. However, promotions contain a significant earnings premium, at least in the short term. Results help to understand inequality in wages and conditions that occur due to sorting, and the acquisition of life chances, like promotions.


# Introduction 


# Job matching and Job searching theories

Two theoretical approaches help to understand job mobility processes, these are the *job-searching* and  *job-matching* approaches. Both suggest that poor work environments, either from inadequate pay or inadequate conditions, push workers to search for new positions [@kalleberg2001; @gensthuizen2008]. The *job searching* approach considers the *Reservation Wage*, or the minimum wage a worker will consider before moving to a new position, given her skill set. *Reservation wage X* drives employers out of unemployment, by capturing the minimum wage that workers will accept before taking a job. *Reservation wage Y* where (X < Y) drives employees to new positions either within or between employers, by capturing the minimum wage that a worker will accept before changing changing positions. If a previously unemployed worker accepts *Wage X* that is below *Wage Y*, they are assumed to continue job searching on-the-job unntil they find their reservation wage [@schmelzer2002; @burdett1978]. The *job-matching* approach is one where workers take time to evaluate whetehr their skills are a good fit for their tasks at work. Since work is an experience good, workers must try several jobs before sticking with a job. This theory would explain why younger workers are typically most likely to change jobs [@sorensen; @thurrow].  

The theories above do not expect a gender difference in patterns of job mobility, at least not explicitly. However, authors routinely find differences  between men and women in their chances of job mobility [@blau1981; @keith; @keith; @keith]. Young women are more likely to quit when comparede to men, and the impact of work tenure is less effective for women than men, with men showing a lower chances of quitting as their time with a given employer increases [@blau1981]. Further, men appear more likely to be rewarded for "economic quits" which are voluntary changes to a new employer, where a respondents cites finding a "better job". This suggests that the wage or conditions premium described above applies more often to men than to women, even when removing gender differences in types of mobility. Some authors argue that the lack of a gender  


# Methodology

This article uses five rounds of Russia's Longitudinal Monitoring Survey (Rounds 20-24), a representative and longitudinal dataset of Russia's population [@kozyreva2016data]. We focus only on those who are in employment and 

# Results

Table one considers gender differences in mobility and earnings, focusing on observations rather than individuals. Overall, respondents are largely immobile although men (76 per cent) appear slightly less likely to list an observation without mobility when compared to women (81 per cent). Both men and women (4 per cent) are equally likely to list a promotion in a given year. However, promotions are generally uncommon. Respondents are much more likely to list exiting a firm, with men (18 per cent) listing more exits than women (12 per cent).

Thinking about earnings, there is an difference between men and women, with immobile observations for men (24,000+ rubles) being higher than immobile observations for women (16,000+ rubles). However, between each group, there appears to be a premium tied to promotion, and a penalty tied to Exit. 

```{r}

ru2 %>% 
  filter(!is.na(mob_final)) %>% 
  group_by(gender, mob_final) %>% 
  summarise(n = n(),
            wage = mean(wage, na.rm=TRUE)) %>% 
  ungroup() %>% 
  group_by(gender) %>% 
  mutate(per = round(n/sum(n)*(100), 2)) %>% 
  ungroup() %>% 
  select(-gender,
         Mobility= mob_final,
         Observations = n,
         Percent = per,
         Wage= wage) %>% 
  kable(format.args = list(decimal.mark = '.', big.mark = ","),
        caption = "Mobility and earnings, by gender") %>% 
  kable_styling(latex_options = 'hold_position') %>% 
  group_rows("Male", 1, 4) %>% 
  group_rows("Female", 5, 8) 



```

Table 2 considers the age differences in these rates. We see that younger groups are more mobile than older groups for both genders, with a seemingly steady increase between respondents aged 0-35, 36-55, and those aged over 55. Young men are still more mobile than young women, but the likelihood of mobility decreases among older respondents.  

```{r}

ru2 %>% 
  filter(!is.na(mob_final)) %>% 
  group_by(gender, age_groups) %>% 
  count(mob_final) %>% 
  mutate(per = round(n/sum(n)*(100), 2)) %>% 
  ungroup() %>% 
  select(Mobility= mob_final,
         Observations = n,
         Percent = per) %>% 
  kable(format.args = list(decimal.mark = '.', big.mark = ","),
        caption = "Mobility, by age and gender") %>% 
  kable_styling(latex_options = 'hold_position') %>% 
  group_rows("Male Aged 0-35", 1, 4) %>% 
  group_rows("Male 36-55", 5, 8) %>% 
  group_rows("Male 56+", 9, 12) %>% 
  group_rows("Female Aged 0-35", 13, 16) %>% 
  group_rows("Female Aged 36-55", 17, 20) %>% 
  group_rows("Female Aged 56+", 21, 24)
```

## Multinomial regression

this section cosiders the likelihood of experiencing a mobility event.

```{r}
d1  <-  ru2 %>% 
  filter(round == 24,
         age_groups=='0-35') 


  
m1  <- multinom(mob_final ~ age + gender + hours + job_year, data = d1)
```


```{r}
tidy(m1, conf.int = T,
     conf.level = 0.95,
     exponentiate = T)


```


# References


